Experience Report: Distributed Computing for the Common Man

### 2010-03-24 doc/icfp-2010/main.tex
*** empty log message ***

The first column shows the value of $N$.
Each computation is divided into a certain number of tasks,
which is shown in second column.
Then the last three column show execution timings in seconds for the
three implementations. The figures within brackets show the speedup
ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks         & sequential& cores         & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00                & 24.90    
    \\\hline
    18 &   18    &           & 123.00               & 125.00     
    \\\hline
       &  272    &           & 103.00               & 124.00
    \\\hline
    19 &   19    &           &                      & 940.00
    \\\hline
       &  306    &           &                      & 819.00      \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} implementation
provides a significant speedup. When the number of tasks is large
enough, the speedup ratio is almost 8, which is also the number of
cores we use. 
% (When the number of tasks is small and not a multiple of
% the number of cores,  for example as in third row with 17 tasks
% executing on 8 cores, the

% TODO comment on network ratios

Where computation cost for each tasks is too small, then communication
cost dominates and could even slow down the execution
w.r.t. sequential implementation, as evident from second row.

@@

The master program proceeds by reading the 80 file names, turning them
into tasks by multiplying them by the number of provers (240 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 600 seconds.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{invalid}, \emph{cannot-decide} (depending on
whether the VC is valid, invalid or undecided by the prover)
and \emph{timeout}.

\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & invalid & cannot & timeout 
    \\\hline\hline
    Alt-ergo & 
    \\\hline
    Simplify & 
    \\\hline
    Z3       & 
    \\\hline
  \end{tabular}
\end{center}

% TODO compare total amount of time wrt sequential time

>>>>

The first column shows the value of $N$.
Each computation is divided into a certain number of tasks,
which is shown in second column.
Then the last three columns show execution timings in seconds for the
three implementations. The figures within brackets show the speedup
ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks  & sequential& cores                 & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
    \\\hline
    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
    \\\hline
       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
    \\\hline
    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
    \\\hline
       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
    \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} and \of{Network}
implementations
provides a significant speedup. As evident from the last row,
the speedup ratio is almost 8, which is also the number of
cores we use. 
It is also evident from the last column that \of{Network}
implementation performs significantly better when the computation time
dominates the total execution time.
The two extreme cases correspond to second and last rows: in the
second row, the communication time dominates and is in fact more than
91\%\ of the total execution time; on the other hand, for the last row
communication time amounts to just 4.6\%\ of the total execution time.
As expected, the network implementation is only beneficial when the
computation time for each individual task is significant, which is the
case in realistic examples.

@@

The master program proceeds by reading the 80 file names, turning them
into tasks by multiplying them by the number of provers (240 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 10 minutes.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Z3       &   4.20 &   0.00 &    160  &  20.0    
    \\\hline
    Alt-ergo &   3.60 &   0.57 &    190  &   8.2     
    \\\hline
    Simplify &   0.02 &   0.00 &     40  &   1.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 7 hours if provers were executed
sequentially. However, using our library and our 3-machines
infrastructure, it completes in 1 hour and 5 minutes, giving us a
speedup of 6.6$\times$.

% timeout = 600
%            |  valid       | invalid      | cannot       | timeout      | failure
% ---------------------------------------------------------------------------------------
%         z3 |  42 (251.25s) |   0 ( 0.00s) |   0 ( 0.00s) |  16 (9548.79s) |  22 (1223.81s) | 
%   alt-ergo |  40 (217.55s) |   0 ( 0.00s) |  19 (34.29s) |  19 (11376.70s) |   2 (489.05s) | 
%   simplify |  32 ( 0.95s) |   0 ( 0.00s) |  34 ( 0.25s) |   4 (2394.60s) |  10 (62.76s) | 
% 0.43user 0.33system 1:04:57elapsed 0%CPU (0avgtext+0avgdata 0maxresident)k

% total amount of time = 25630 seconds = 7 hours 7 minutes 10 seconds
% ratio = 6.57

% timeout = 120
%            |  valid       | invalid      | cannot       | timeout      | failure
% ---------------------------------------------------------------------------------------
%         z3 |  41 (211.08s) |   0 ( 0.00s) |   0 ( 0.00s) |  20 (2383.89s) |  19 (262.30s) | 
%   alt-ergo |  40 (237.00s) |   0 ( 0.00s) |  19 (34.45s) |  21 (2514.42s) |   0 ( 0.00s) | 
%   simplify |  32 ( 0.98s) |   0 ( 0.00s) |  34 ( 0.31s) |   5 (598.92s) |   9 (18.98s) |
